UDC 911.3

### 2024-02-12 maksakov-readings-conf-2024/Conference-paper.Rmd
Minor fixes in CSL translation and docx styles, compression of paper text, finish of translation of plots and tables, fix of a bug with paths table

---
title: "Migrations of Russian Commercial Legal Companies in 2016–23"
author: P. O. Syomin
output: 
  word_document:
    reference_docx: conference-paper-reference.docx
bibliography: refs.bib
csl: "gost-r-7-0-5-2008-numeric-alphabetical.csl"
---

@@

Migrations of Russian legal companies in 2016–23 are described and analysed. The study uses Federal Tax Service open data on small and medium-sized companies. Distributions by migration count and distance are estimated and visualized, the relationship between company size and migration decision is examined and statistically tested, net migration rates by region are mapped, and top migration paths are listed. It is found that migrations are rare, mostly short-distance (~10 km). Bigger companies tend to relocate more often. Moscow and St. Petersburg have probably lost small and inefficient businesses but gained larger and more profitable.

@@

*Keywords:* companies migration, small and medium-sized businesses, legal companies, legal service, service geography, data analysis, open data, Federal Tax Service.

@@

Migration of companies, also known as business relocation, is an important phenomenon of economics and geography. Its first studies date back to the middle of 20th century, starting from G. McLaughlin and S. Robock's book *Why industry moves South* that have described the migration of enterprises from the North of the US to the South and linked it to the cheaper labor and smaller activity of labor unions [@McLaughlin1949]. Business relocation may be seen from spatial, relational or decision-making point of view [@Carrincazeaux2015]. A vast amount of research is devoted to various aspects of companies migration, including factors (reasons) [@Nakosteen1987; @Wu2007], directions (places) [@Rupasingha2018; @Rossi2019], discourses [@Piotti2007], taxation effects [@Kubicova2016]. In Russia, there are few papers about migration of companies. Their authors focus mainly on the issue of relocation as an instrument of tax optimization, thus discussing only “negative” part of the problem [@Adigamova2011_en; @Shcherbakova2016_en] without diving into factors or places and mostly skipping the geographical aspects of the phenomenon. Thus, it appears that a simple spatial analysis of business relocation in Russia is important. The present study addresses this issue by describing and visualizing the migration of Russian commercial legal companies during the period from the middle of 2016 to the end of 2023. The selection of concrete area of services, i.e. legal firms, is quite a subjective choice, but looks like it is relatively representative. Moreover, it is likely that legal business is more mobile and more sensitive to non-economic factors, although it is just a hypothesis.

**Data.** For the analysis, a complete geo-referenced dataset of Russian small and medium-sized enterprises (SMEs) was used. It was generated from the open and freely available dumps of SME registry distributed by Federal Tax Service (FTS) of Russia [^1]. For the generation of the dataset, the author have developed an open source Python CLI application *ru-smb-companies* [^2]. To select legal companies only, the filtering by activity code 69.10 (*activity in the area of law* according to the All-Russia Activity Code Classifier) was used. The dataset covers the period from August 2016 to November 2023. It contains information about `r companies_count` businesses in tabular format. The data includes, in particular, the following fields: name of the company, its registration address, coordinates of the registration settlement, revenue, expenditure, count of employees (the three last entries are available for each year from 2019 to 2022). It is apperant that some legal companies are missing in the dataset, e.g. large firms which exceeds the limit on the revenue or number of employees for SMEs or advocacy organizations which are categorized as non-profit companies according to the Russian laws and so cannot be included to the list of SMEs. However, there are high chances that it nevertheless can be used as a relevant data source. According to the analysis by Instutite for the Rule of Law [@Moiseeva2016_en], by August 2016 in Russia there have been 47,000 commercial legal companies, and the dataset used here lists `r `round(companies_count_08_2016, 0)` companies for August 2016. So, we can conclude with high confidence that 3/4 of Russian commercial legal companies are present in our dataset, and this fact proves its reliability.

@@

**Methods.** Here, the following definition of migration of company is used: migration is a change in address of registration [@Shcherbakova2016_en]. Maybe it is too narrow, because does not account for other ways of relocation, e. g. mergers or reorganization. Also, it can be criticized as formalistic, because the change of nominal address does not necessarily reflects the change of place of doing business. However, it looks the best choice for the data available, because the dataset does not contain information about mergers and reorganizations, and does not tell us anything about the actual location of business activity. The analysis is done by the means of descriptive statistics, statistical tests, data visualization and mapping. It includes five elements: (1) distribution of legal companies by the count of migrations, (2) relationship between the migration decision and size of a company; (3) distribution of migration distance; (4) net migration rate by region; (5) net migration rate by inter-region migration paths.

To quantitatively estimate the migration, the first or all the three of the following variables are used: count of companies, sum of revenue, sum of count of employees. Technically, the analysis was performed with R programming language (including additional packages). The source code for the analysis is stored in public repository [^3].

@@

A **distribution of companies by the number of migrations** is shown on *Fig. 1* (see below). Almost all the companies (97%) in 2016–2023 have not migrated. About 2.5% of companies have changed the location exactly once, and more frequent migrations are extremely rare. It is seen that the distribution is almost log10-linear: the count in each next group is ~10 times smaller.

@@

```{r fig-1-mig-distr, fig.cap="*Fig. 1.* The distribution of companies by the number of migrations. Y axis is log10-scaled. The counts are shown above the bars, and percetrages are below the top border", fig.height=3}
mig_count_distr_plot
```

@@

A **relationship between the migration choice and the size of company** is depicted on Fig. 2. Migration among small firms is more frequent in comparison to mircocompanies: according to exact Fisher test, `r ifelse(categories_test$p.value < 0.001, "p < 0.001", paste("p = ", categories_test$p.value))`, and odds ratio is `r paste0("1 to ", round(1 / categories_test$estimate, 0))`.

@@

```{r fig-2-mig-by-company-size, fig.cap="*Fig. 2*. Migration choice depending on company's size (category)", fig.height=2}
mig_by_category_plot
```

@@

A **distribution of migration distance** is shown on *Fig. 3*. The majority of businesses moves to small distance (about 10 km), but there is a significant group of long-migrating firms (the distance is about 1000 km).

@@

```{r fig-3-distance, fig.cap="*Fig. 3.* Distribution of migration distance", fig.height=3}
mig_distance_plot
```

@@

**Net migration rate by region** is shown on map (see *Fig. 4*). Apparently, most of the regions have net migration rate close to zero. However, there are some regions with substantial growth or decline in the number of registered legal companies. For example, Moscow region is the top by growth, and the moderate positive net migration rate is also characteristic of Saint Petersburg, Krasnodar, Rostov, Novosibirsk regions (oblasts), Perm region (krai) and Udmurt republic. The desrease in the number of legal businesses is typical for Volga region, the Urals, Western Siberia and Krasnoyarsk region (krai). It is worth mentioning that Moscow itself (as a city) also shows the decline. However, it is also obvious that the absolute numbers of growth or decline are small compared to the absolute number of registered companies and do not exceed 100 in any of the regions.

@@

```{r fig-4-mig-by-region-plot, fig.cap="*Fig. 4.* Net migration rate of legal companies by regions", fig.height=4}
mig_by_region_plot
```

@@

**The most popular migration paths** (found by the number of migrated companies) with corresponding total revenues and employee counts are listed in the **Table**. It is important to note that the first two of them represent the moving from two largest cities of Russia (including the capital) to the surrounding regions (oblasts). However, the migration from other regions to the capital is popular too. Besides, the increase in the number of registered businesses is not necessarily associated with the increase in income or labor force, and vise versa, that is, even if a net migration rate is above zero, the total revenue and employees count may have decreased, possibly meaning that arriving businesses are profitable and large, while leaving companies are small and inefficient. This is the case, for example, of the aforementioned Moscow-to-region and Saint-Petersburg-to-region paths, where the net change of the number of firms is in favor of regions, while the net change in revenue and labor is positive for the cities.

@@

```{r table}
kable(
  head(regional_paths_combined, 10),
  col.names = c("Region", "Count", "Revenue", "Employees"),
  caption = "Top 10 paths of inter-regional migration of Russian legal companies in 2016–2023 and their features: net count of relocated businesses, net total revenue, net total number of employees"
)
```

**Conclusions.** Migration of legal companies in Russia is a rare phenomenon: just about 2.5% of all businesses in this activity area have changed their nominal location in the period of 2016–23. Small companies, in comparison to micro-businesses, are more likely to relocate: the chances are 4 times higher. Short-distance (~10 km) migrations are more common than long-distance (100–1000 km) ones. Moscow and St. Petersburg regions are at the top by net migration rate estimated by the count of relocated businesses, while Moscow have lost the highest number of companies. However, it looks like that while losing small and inefficient companies, Moscow and St. Patersburg attract more profitable businesses.

>>>>

---
title: "UDC 911.3"
subtitle: "Migration of Russian Legal Companies in 2016–23"
author: P. O. Syomin
output: 
  word_document:
    reference_docx: conference-paper-reference.docx
bibliography: refs.bib
csl: "gost-r-7-0-5-2008-numeric-alphabetical-en.csl"
---

@@

Migration of Russian commercial legal companies in 2016–23 is described and analysed. Distributions by migration count and distance are estimated and visualized, the relationship between company size and migration decision is statistically tested, net migration rates by region are mapped, and top migration paths are listed. It is found that migrations are rare, mostly short-distance (10–50 km). Bigger companies tend to relocate more often. Moscow has probably lost small and inefficient businesses but gained larger and more profitable.

@@

***Keywords:*** companies migration, small and medium-sized businesses, legal companies, legal service, service geography, data analysis, open data, Federal Tax Service.

@@

Migration of companies (business relocation) is an important phenomenon of economy and geography. Its first studies date back to the middle of 20th century, starting from G. McLaughlin and S. Robock's book *Why industry moves South* that have described the migration of enterprises from the North of the US and linked it to the cheaper labor and smaller activity of labor unions [@McLaughlin1949]. Various aspects of companies migration have been studied, including factors (reasons) [@Wu2007], directions (places) [@Rossi2019], taxation effects [@Kubicova2016].

In Russia, there are few papers about migration of companies. Existing works focus  on relocation as an instrument of tax optimization [@Adigamova2011_en; @Shcherbakova2016_en], skipping the geographical overview and spatial aspects of the phenomenon. So, a spatial analysis of business relocation in Russia is important. This study describes and visualizes the migration of Russian commercial legal companies in 2016–23. The choice of particular area of economy, that is, legal service, is subjective.

**Data.** A geo-referenced dataset of Russian small and medium-sized businesses (SMBs) with main activity code equal to 69.10 (activity in the area of law) was generated from the open dumps of SMB registry[^1] with an open source Python CLI application *ru-smb-companies*[^2] developer by the author of this paper. The dataset covers the period from August 2016 to November 2023. According to [@Moiseeva2016_en], by August 2016 in Russia there have been 47000 commercial legal companies, and the dataset contains about `r round(companies_count_08_2016, -3)` companies with the same date. Some legal companies are missing, e. g. large ones that exceed the limit on the revenue or number of employees for SMBs, but 3/4 are present, so the dataset looks reliable and relevant.

@@

**Methods.** Company migration is seen as a change in address of registration [@Shcherbakova2016_en]. It is narrow and does not account for other ways of relocation (e. g. mergers or reorganization). Also, it is formal, because the change of nominal address does not necessarily reflects the change of actual place of doing business. However, it looks like the best choice for the data available. Descriptive statistics, statistical tests, data visualization and mapping are used. Technically, the analysis was performed with R programming language (including additional packages). The source code for the analysis is stored in public repository[^3].

@@

A **distribution of companies by the number of migrations** is shown on *Fig. 1*. Almost all the companies (97%) in 2016–23 have not migrated. About 2.5% of companies have changed the location exactly once, and more frequent migrations are extremely rare. The distribution is almost log10-linear.

@@

```{r fig-1-mig-distr, fig.cap="*Fig. 1.* The distribution of companies by the number of migrations. Y axis is log10-scaled. The counts are shown above the bars, and percetrages are below the top border", fig.height=2.5}
mig_count_distr_plot
```

@@

A **relationship between the migration decision and the size of company** is depicted on *Fig. 2*. Migration among small firms is more frequent in comparison to micro-businesses: according to Fisher's exact test, `r ifelse(categories_test$p.value < 0.001, "p < 0.001", paste("p = ", categories_test$p.value))`, and odds ratio is `r paste0("1 to ", round(1 / categories_test$estimate, 0))`.

@@

```{r fig-2-mig-by-company-size, fig.cap="*Fig. 2*. Migration decision depending on company's size (category)", fig.height=1}
mig_by_category_plot
```

@@

A **distribution of migration distance** is shown on *Fig. 3*. The majority of businesses moves to small distance (10–50 km), but there is a significant group of long-migrating firms (the distance is about 1000 km).

@@

```{r fig-3-distance, fig.cap="*Fig. 3.* Distribution of migration distance", fig.height=2.5}
mig_distance_plot
```

@@

**Net migration rate by region** is shown on map (see *Fig. 4*). Most of the regions have net migration rate close to zero. However, there are some regions with substantial growth (Moscow, St. Petersburg, Krasnodar, Rostov, Novosibirsk oblast, Perm krai) or decline (Moscow as a city) in the number of registered legal companies. The absolute counts are small (<100) compared to the quantity of registered companies.

@@

```{r fig-4-mig-by-region-plot, fig.cap="*Fig. 4.* Net migration rate of legal companies by regions", fig.height=3}
mig_by_region_plot
```

@@

**The most popular migration paths** (by the count of migrated companies) with summed up revenues and employee counts are listed in the **Table**. The first two paths represent the relocation from two largest cities of Russia to the surrounding regions. Net change in the number of businesses is relatively small. An increase in the number of registered businesses is not necessarily associated with the increase in income or labor force, and vise versa. Positive flow of companies with negative flow of revenue or labor probably means that arriving businesses are larger and more profitable than leaving.

@@

```{r table}
kable(
  head(regional_paths_combined_en, 10),
  col.names = c("Path", "Count", "Revenue, M rub", "Employees"),
  caption = "Top 10 paths of inter-regional migration of Russian legal companies in 2016–23",
  digits = 0
)
```

**Conclusion.** Migration of legal companies in Russia is a rare phenomenon: only about 2.5% of all businesses in this activity area have changed their nominal location in the period of 2016–23. Small companies, in comparison to micro-businesses, are more likely to relocate: the chances are 4 times higher. Short-distance (10–50 km) migrations are more common than long-distance (~1000 km) ones. Moscow and St. Petersburg regions are at the top by net migration rate estimated by the count of relocated businesses, while Moscow have lost the highest number of companies. However, it looks like that while losing small and inefficient companies, Moscow attracts more profitable businesses.
